### 2020-10-08 analysis/analysis.Rmd
starting to clean up data

---
title: "Dragonfly Territory Ananlysis"
Author: Maggie Vincent
Term: Fall 2020
Course: Reproducible Research in R
output: pdf_document
---

# Read in Data from BioTime

rawDragonfly <- `rawdata.(1)`

install.packages('plyr')
install.packages('dplyr')

library('plyr')
library('dplyr')

coenigrion <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
ischnura <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
lestes <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
libelsympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
sympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')