### 2020-10-29 analysis/analysis.Rmd
subsetting data.frames

---
title: "Dragonfly Territory Ananlysis"
Author: "Maggie Vincent"
Term: Fall 2020
Course: Reproducible Research in R
output: pdf_document
---

# Read in Data from BioTime

rawDragonfly <- `rawdata.(1)`

install.packages('plyr')
install.packages('dplyr')

library('plyr')
library('dplyr')

coenigrion <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
ischnura <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
lestes <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
libelsympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
sympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')

# Make a rough plot to visualize the data

par(mfrow=c(2,3))
plot(ischnura$YEAR, ischnura$ABUNDANCE)
plot(coenigrion$YEAR, coenigrion$ABUNDANCE)
plot(lestes$YEAR, lestes$ABUNDANCE)
plot(libellula$YEAR, libellula$ABUNDANCE)
plot(sympetrum$YEAR, sympetrum$ABUNDANCE)

# I don't like that the y-axes are scaled differently and I want to change the axis labels

# subsetting down the data.frames, since I need to do it more than once, I'll write a simple function

getAbundanceYear <- function(dat, x = 'ABUNDANCE', y = 'YEAR'){
                    tmp <- subset(dat, select = c(x, y))
                    return(tmp)
}

ischnuraAY <- getAbundanceYear(ischnura)
coenigrionAY <- getAbundanceYear(coenigrion)
lestesAY <- getAbundanceYear(lestes)
libellulaAY <- getAbundanceYear(libellula)
sympetrumAY <- getAbundanceYear(sympetrum)

# the current data.frames are different lengths
# this is because the years without observations of the species were not put in as zeroes
# after consultation with Tad Dallas, I will be changing the data.frames to length 27 (observation years) with zeroes for the missing years