### 2020-11-13 analysis/analysisSS.Rmd

---
title: ""
Author: "Maggie Vincent"
Term: Fall 2020
Course: Reproducible Research in R
output: pdf_document
---

# Read in Data from BioTime - Single study
## this was downloaded from 'http://biotime.st-andrews.ac.uk/taxa.php?10' on 4 October 2020
## study name == 'The development of dragonfly communities and the consequences of territorial behaviour'
## study ID == ''
## BioTIME does not have an API, making this necessary

rawDragonfly <- `DataSingleStudy`

install.packages('plyr')
install.packages('dplyr')

library('plyr')
library('dplyr')

coenigrion <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
ischnura <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
lestes <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
libellula <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
sympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')

# Make a rough plot to visualize the data

par(mfrow=c(2,3))
plot(ischnura$YEAR, ischnura$ABUNDANCE)
plot(coenigrion$YEAR, coenigrion$ABUNDANCE)
plot(lestes$YEAR, lestes$ABUNDANCE)
plot(libellula$YEAR, libellula$ABUNDANCE)
plot(sympetrum$YEAR, sympetrum$ABUNDANCE)

# I don't like that the y-axes are scaled differently and I want to change the axis labels

# subsetting down the data.frames, since I need to do it more than once, I'll write a simple function

getAbundanceYear <- function(dat, x = 'ABUNDANCE', y = 'YEAR'){
                    tmp <- subset(dat, select = c(x, y))
                    return(tmp)
}

ischnuraAY <- getAbundanceYear(ischnura)
coenigrionAY <- getAbundanceYear(coenigrion)
lestesAY <- getAbundanceYear(lestes)
libellulaAY <- getAbundanceYear(libellula)
sympetrumAY <- getAbundanceYear(sympetrum)

# the current data.frames are different lengths
# this is because the years without observations of the species were not put in as zeroes
# after consultation with Tad Dallas, I will be changing the data.frames to length 27 (observation years) with zeroes for the missing years

row1959 <- c(0, 1959)
row1960 <- c(0, 1960)
row1961 <- c(0, 1961)
row1962 <- c(0, 1962)
row1963 <- c(0, 1963)

row1972 <- c(0, 1972)
row1974 <- c(0, 1974)

coenigronCompAY <- rbind(row1959, row1960, row1961, row1962, coenigrionAY)
ischnuraCompAY <- rbind(row1959, row1960, row1961, ischnuraAY)
lestesCompAY <- rbind(row1959, row1960, row1961, row1962, row1963, lestesAY)
# libellula will be dealt with differently because it is missing rows within the dataframe
sympetrumCompAY <- rbind(row1959, row1960, row1961, sympetrumAY)

prelibellulaCompAY <- rbind(row1960, row1972, row1974, libellulaAY)
# this will add the rows I want, but now in chronological order
libellulaCompAY <- dplyr::arrange(prelibellulaCompAY, YEAR)

# there are species missing from this, they need to be added separately

brachytronAbundance <- c(0,0,0,0,0,0,0,0,0,1,0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
pyrrhosomaAbundance <- c(0,0,3,20,18,5,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
aeshnaAbundance <- c(0,0,0,0,1,2,3,1,1,2,3,1,5,1,2,0,0,0,1,1,1,0,2,1,1,3,1,2,2,2)