### 2020-12-02 manuscript/manuscript.Rmd
working on results and figures in manuscript

---
title: "Odonate abundance changes in response to the creation of a flood reservoir in a former fen"
author: Maggie Vincent
output:
  pdf_document:
    fig_caption: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
install.packages('plyr')
install.packages('dplyr')
```

# Introduction

The invertebrate order Odonata is comprised of at least 5680 freshwater insects and are commonly referred to as damselflies and dragonflies. Due to the larvae’s need for freshwater, they can be used as gauges for local environmental quality and health (Kalkman *et al.* 2008). Destruction of habitation and changes in water quality have led to decreases in species distributions, although extinction is not an immediate worry for most, as they often have broad distributions (Kalkman *et al.* 2008).

The order is of particular interest for their reproductive behaviors, with males of the species utilizing the aedeagus not only for sperm transfer from a secondary sexual organ, but also for the removal of sperm already in a female’s storage organs (Waage, 1979). Additionally, because many species are quite conspicuous in their coloration, they are possible for identification by non-experts. This makes them an interesting and entertaining aspect of outdoor recreation and wildlife hobbyists.

Although the highest biodiversity of Odonates resides in the Neotropical and Oriental regions (Kalkman *et al.* 2008), Odonates extend into temperate zones, albeit with less diversification. This project is concerned with East Anglia, a region in eastern England. This area was formerly swampy, or fenland, until it was drained for agricultural use over the last few centuries (DeFries *et al.* 2004). Habitat changes such as fragmentation and/or loss often lead to decreases in biodiversity (DeFries *et al.* 2004), as stated above for Odonates. Conservation and preservation attempts over the last century have moved to protect remaining areas and return others back to their original (or similar) functions. There has been work since the late 1950s in East Anglia to preserve ancient fens with the possibility of reconnecting the remaining fragments into a “Great Fen” (“Great”, n.d.) in the relevant area.

The recent movement to assess the environmental-agricultural trade-offs of reclaimed lands in England (and elsewhere) requires a careful balance between the environmental benefits of conservation and the financial difficulties for farmers and locals (Morris et al. 2000). There is the possibility of determining areas of lower agricultural production that can be returned to their former state without detrimental effects to local famers (Morris *et al.* 2000; “Great”, n.d.).

# Objectives

The goals of this study can be separated into two main purposes: description of changing (or not) species abundance over more than two decades and a focus on transparency and reproducibility in the workflow.

* Does *Odonata* abundance significantly change over the course of the study?
* Do species appear to follow different patterns of abundance changes?
* Do certain species seem to suffer upon the introduction of others?
  + How does this match up with current citizen science?

The first three main objectives will be answered from the dataset from the original study (Morris, 1991). The final sub-bullet is not meant to be answered with true rigor, but as an interesting way to include citizen science. It is not expected to be comprehensive or a full picture of current Odonate diversity at the study site.

## Data Source

The data for this descriptive study is available through BioTIME (biotime.st-andrews.ac.uk/downloadArea.php), a database devoted to biodiversity. The database is dynamic and does not have an API, making reproducibility via the database a step more difficult. The dataset itself is constant, but its online source changing. The dataset was not originally representative of the first study’s findings, and required significant manipulation. The methods and resulting implications for reproducibility will be discussed in more detail in later sections.

```{r, echo=FALSE, fig.cap = 'Sample of the unmanipulated data set.'}
DataSingleStudy <- read.csv("~/GitHub/rrirfinalproject/DataSingleStudy")
head(DataSingleStudy)
```

## Study Site

The work was carried out at Woodwalton Fen (Long, Lat) in Huntingdon, England, from 1959 to 1988. The area was primarily an ancient fen, or a wetland, until most of the area was drained in the 1800s for agricultural use. In the mid-1900s the study area was restructured as a flood reservoir, recreating the wetland aspect. The original study followed Odonata species to characterize territorial changes and patterns throughout the project (Citation).

![**Figure 1.** Map of eastern England with Woodwalton Fen marked.](Woodwalton_Fen_England.jpg)

## Approach

The original study characterized the return and territoriality of dragonflies and damselflies at the site. In this study, I seek to better characterize the significance of species-specific changes over the study period in order to tease apart any possible changes in overall abundance. The main questions were outlined earlier in the objectives.

# Methods

The data used is publicly available for download and use from BioTIME (Dornela *et al.* 2018) or from the Github repository (https://github.com/mvcarneal/rrirfinalproject) hosting the analytical pipeline for this project. The data was manipulated (with an original retained) in R Version 4.0.2 with the additional packages ‘plyr’ and ‘dplyr’. The manipulated dataset is also available on the aforementioned Github repository, along with the code used in the analysis folder. As the dataset was previously compiled and published, the methods for this project will concern the approach taken and code utilized.

Base, plyr and dplyr were utilized in an attempt to keep the number of packages small and the threshold for reproducibility lower, and Docker used to create a container for reproducibility. The latter two aided in the manipulation of the dataset.

# Results

```{r, echo=FALSE}
library('plyr')
library('dplyr')

DataSingleStudy <- read.csv("~/GitHub/rrirfinalproject/DataSingleStudy")
rawDragonfly <- `DataSingleStudy`

cPuella <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
iElegans <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
lSponsa <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
lQuadrimaculata <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
sStriolatum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')

getAbundanceYear <- function(dat, x = 'ABUNDANCE', y = 'YEAR'){
                    tmp <- subset(dat, select = c(x, y))
                    return(tmp)
}

iElegansAY <- getAbundanceYear(iElegans)
cPuellaAY <- getAbundanceYear(cPuella)
lSponsaAY <- getAbundanceYear(lSponsa)
lQuadrimaculataAY <- getAbundanceYear(lQuadrimaculata)
sStriolatumAY <- getAbundanceYear(sStriolatum)

iElegansSAY <- dplyr::rename(iElegansAY, iElegansAbundance = ABUNDANCE)
cPuellaSAY <- dplyr::rename(cPuellaAY, cPuellaAbundance = ABUNDANCE)
lSponsaSAY <- dplyr::rename(lSponsaAY, lSponsaAbundance = ABUNDANCE)
lQuadrimaculataSAY <- dplyr::rename(lQuadrimaculataAY, lQuadrimaculataAbundance = ABUNDANCE)
sStriolatumSAY <- dplyr::rename(sStriolatumAY, sStriolatumAbundance = ABUNDANCE)

row1959 <- c(0, 1959)
row1960 <- c(0, 1960)
row1961 <- c(0, 1961)
row1962 <- c(0, 1962)
row1963 <- c(0, 1963)

row1972 <- c(0, 1972)
row1974 <- c(0, 1974)

cPuellaCompAY <- rbind(row1959, row1960, row1961, row1962, cPuellaSAY)
iElegansCompAY <- rbind(row1959, row1960, row1961, iElegansSAY)
lSponsaCompAY <- rbind(row1959, row1960, row1961, row1962, row1963, lSponsaSAY)
# lQuadrimaculata will be dealt with differently because it is missing rows within the dataframe
sStriolatumCompAY <- rbind(row1959, row1960, row1961, sStriolatumSAY)

prelibellulaCompAY <- rbind(row1960, row1972, row1974, lQuadrimaculataSAY)
# this will add the rows I want, but now in chronological order
lQuadrimaculataCompAY <- dplyr::arrange(prelibellulaCompAY, YEAR)

# there are species missing from this dataset, they need to be added separately, I will source (and cite) the abundance data
# also need a vector corresponding to the years

YEAR <- c(1959:1988)

bPratenseAbundance <- c(0,0,0,0,0,0,0,0,0,1,0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
pNymphulaAbundance <- c(0,0,3,20,18,5,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
aCyaneaAbundance <- c(0,0,0,0,1,2,3,1,1,2,3,1,5,1,2,0,0,0,1,1,1,0,2,1,1,3,1,2,2,2)
sSanguineumAbundance <- c(0,0,0,0,0,1,0,0,1,0,2,0,1,0,0,1,0,1,4, 11, 3, 0, 1, 9, 6, 10, 4, 9, 3, 4)
aGrandisAbundance <- c(0,0,0,0,0,0,0,4,0,1,1,0,2,0,2,0,0,0,1,1,0,1,2,1,0,1,1,4,0,1)

bPratenseCompAY <- data.frame(bPratenseAbundance, YEAR)
pNymphulaCompAY <- data.frame(pNymphulaAbundance, YEAR)
aCyaneaCompAY <- data.frame(aCyaneaAbundance, YEAR)
sSanguineumCompAY <- data.frame(sSanguineumAbundance, YEAR)
aGrandisCompAY <- data.frame(aGrandisAbundance, YEAR)

# I am going to join these 10 dataframes by the common column year
# I want year on the far left so I am going to reorder the first dataframe that will determine the order of my join function

bPratenseCompAY2 <- subset(bPratenseCompAY, select = c(2,1))

join1 <- dplyr::left_join(bPratenseCompAY2, pNymphulaCompAY, by= NULL)
join2 <- dplyr::left_join(join1, aCyaneaCompAY, by = NULL)
join3 <- dplyr::left_join(join2, sSanguineumCompAY, by = NULL)
join4 <- dplyr::left_join(join3, aGrandisCompAY, by = NULL)
join5 <- dplyr::left_join(join4, cPuellaCompAY, by = NULL)
join6 <- dplyr::left_join(join5, iElegansCompAY, by = NULL)
join7 <- dplyr::left_join(join6, lSponsaCompAY, by = NULL)
join8 <- dplyr::left_join(join7, lQuadrimaculataCompAY, by = NULL)
fullAbundanceDF <- dplyr::left_join(join8, sStriolatumCompAY, by = NULL)

# I want a total for each year

abundanceSums <- data.frame(rowSums(fullAbundanceDF[,c(2,3,4,5,6,7,8,9,10,11)]))
abundance_year <- cbind(abundanceSums, new_col = YEAR)
abundanceYearSums <- subset(abundance_year, select = c(2,1))

abunYearSums_1 <- dplyr::rename(abundanceYearSums, year = new_col)
abunYearsSums <- dplyr::rename(abunYearSums_1, total_abundance = rowSums.fullAbundanceDF...c.2..3..4..5..6..7..8..9..10..11...)
```

The damselflies and dragonflies in this study are listed in Table 1 for reference. The figure below (Fig. 2), is a linear regression model of the maximum abundance of males in a single day from 1959  to 1988. The abundance corresponds to the sum of the ten species in Table 1 and the p-value for the model was 0.000162.

```{r, echo=FALSE,}

# weird character in genus column name, need to rename it, dplyr function

odonata_Woodwalton <- dplyr::rename(odonata_names, GENUS = ï..GENUS)

knitr::kable(odonata_Woodwalton, caption = 'Table 1. Species and common names of ten Odonates found at Woodwalton Fen between 1959 and 1988.')

```

```{r, echo=FALSE, fig.align=center, fig.cap='**Figure 2** Linear model of abundance of all ten species from 1959 to 1988. p-value = 0.000162 }
par(mfrow=c(1,1))
plot(abunYearsSums$year, abunYearsSums$total_abundance, xlab = "Year", ylab = "Abundance")
abline(lm(abunYearsSums$total_abundance ~ abunYearsSums$year), col="red")
```

```{r, echo=TRUE}
fitTotal <- lm(abunYearsSums$total_abundance ~ abunYearsSums$year)

summary(fitTotal)

pValueTotal <- 0.000162

```

```{r, echo=FALSE, fig.cap=paste("Figure 3. Abundances broken down by species with line of best fit.") }
par(mfrow=c(2,5))
plot(bPratenseCompAY$YEAR, bPratenseCompAY$bPratenseAbundance, main = "Brachytron pratense", xlab = "Year", ylab = "Abundance")
abline(lm (bPratenseCompAY$bPratenseAbundance ~ bPratenseCompAY$YEAR), col="red")

plot(pNymphulaCompAY$YEAR, pNymphulaCompAY$pNymphulaAbundance, main = "Pyrrhosa nymphula", xlab = "Year", ylab = "Abundance")
abline(lm (pNymphulaCompAY$pNymphulaAbundance ~ pNymphulaCompAY$YEAR), col="red")

plot(aCyaneaCompAY$YEAR, aCyaneaCompAY$aCyaneaAbundance, main = "Aeshna cyanea", xlab = "Year", ylab = "Abundance")
abline(lm(aCyaneaCompAY$aCyaneaAbundance ~ aCyaneaCompAY$YEAR), col="red")

plot(sSanguineumCompAY$YEAR, sSanguineumCompAY$sSanguineumAbundance, main = "Sympetrum sanguineum", xlab = "Year", ylab = "Abundance")
abline(lm(sSanguineumCompAY$sSanguineumAbundance ~ sSanguineumCompAY$YEAR), col="red")

plot(aGrandisCompAY$YEAR, aGrandisCompAY$aGrandisAbundance, main = "Aeshna grandis", xlab = "Year", ylab = "Abundance")
abline(lm(aGrandisCompAY$aGrandisAbundance ~ aGrandisCompAY$YEAR), col="red")

plot(cPuellaCompAY$YEAR, cPuellaCompAY$cPuellaAbundance, main = "Coenagrion puella", xlab = "Year", ylab = "Abundance")
abline(lm(cPuellaCompAY$cPuellaAbundance ~ cPuellaCompAY$YEAR), col="red")

plot(iElegansCompAY$YEAR, iElegansCompAY$iElegansAbundance, main = "Ischnura elegans", xlab = "Year", ylab = "Abundance")
abline(lm(iElegansCompAY$iElegansAbundance ~ iElegansCompAY$YEAR), col="red")

plot(lSponsaCompAY$YEAR, lSponsaCompAY$lSponsaAbundance, main = "Lestes sponsa", xlab = "Year", ylab = "Abundance")
abline(lm(lSponsaCompAY$lSponsaAbundance ~ lSponsaCompAY$YEAR), col="red")

plot(lQuadrimaculataCompAY$YEAR, lQuadrimaculataCompAY$lQuadrimaculataAbundance, main = "Libellula quadrimaculata", xlab = "Year", ylab = "Abundance")
abline(lm(lQuadrimaculataCompAY$lQuadrimaculataAbundance ~ lQuadrimaculataCompAY$YEAR), col="red")

plot(sStriolatumCompAY$YEAR, sStriolatumCompAY$sStriolatumAbundance, main = "Sympetrum striolatum", xlab = "Year", ylab = "Abundance")
abline(lm(sStriolatumCompAY$sStriolatumAbundance ~ sStriolatumCompAY$YEAR), col="red")

```

```{r, echo=FALSE}
library('plyr')
library('dplyr')

names_p_values <- read.csv("~/GitHub/rrirfinalproject/analysis/names_p_values.csv")

# strange character in column name for species

corrected_names_p_values <- dplyr::rename(names_p_values, Species = ï..SPECIES)

knitr::kable(corrected_names_p_values, caption = 'Table 2. p-values from the LM for each species and all species combined shown in Figure 2. Significance shown with asterisks.')

```

```{r, echo=FALSE, fig.cap="Figure 4. Number of species in each year of the study period. Note that this is not a comprehensive list of all Odonates at the site, this will be addressed in the discussion."}
# I want to know how many species there are in a given year, subsetting using the which function and finding the length of that subset
# I will save each as with a corresponding year identifier because I have not yet decided what format I want them all in (dataframe, table, etc.)

ASpecies <- length(which(fullAbundanceDF[1,2:11] != 0))
BSpecies <- length(which(fullAbundanceDF[2,2:11] != 0))
CSpecies <- length(which(fullAbundanceDF[3,2:11] != 0))
DSpecies <- length(which(fullAbundanceDF[4,2:11] != 0))
ESpecies <- length(which(fullAbundanceDF[5,2:11] != 0))
FSpecies <- length(which(fullAbundanceDF[6,2:11] != 0))
GSpecies <- length(which(fullAbundanceDF[7,2:11] != 0))
HSpecies <- length(which(fullAbundanceDF[8,2:11] != 0))
ISpecies <- length(which(fullAbundanceDF[9,2:11] != 0))
JSpecies <- length(which(fullAbundanceDF[10,2:11] != 0))
KSpecies <- length(which(fullAbundanceDF[11,2:11] != 0))
LSpecies <- length(which(fullAbundanceDF[12,2:11] != 0))
MSpecies <- length(which(fullAbundanceDF[13,2:11] != 0))
NSpecies <- length(which(fullAbundanceDF[14,2:11] != 0))
OSpecies <- length(which(fullAbundanceDF[15,2:11] != 0))
PSpecies <- length(which(fullAbundanceDF[16,2:11] != 0))
QSpecies <- length(which(fullAbundanceDF[17,2:11] != 0))
RSpecies <- length(which(fullAbundanceDF[18,2:11] != 0))
SSpecies <- length(which(fullAbundanceDF[19,2:11] != 0))
TSpecies <- length(which(fullAbundanceDF[20,2:11] != 0))
USpecies <- length(which(fullAbundanceDF[21,2:11] != 0))
VSpecies <- length(which(fullAbundanceDF[22,2:11] != 0))
XSpecies <- length(which(fullAbundanceDF[23,2:11] != 0))
YSpecies <- length(which(fullAbundanceDF[24,2:11] != 0))
ZSpecies <- length(which(fullAbundanceDF[25,2:11] != 0))
AASpecies <- length(which(fullAbundanceDF[26,2:11] != 0))
ABSpecies <- length(which(fullAbundanceDF[27,2:11] != 0))
ACSpecies <- length(which(fullAbundanceDF[28,2:11] != 0))
ADSpecies <- length(which(fullAbundanceDF[29,2:11] != 0))
AESpecies <- length(which(fullAbundanceDF[30,2:11] != 0))

# number of species in chronological order
# I'll reuse the years vector I created earlier to make a data.frame

speciesChrono <- c(ASpecies, BSpecies, CSpecies, DSpecies, ESpecies, FSpecies, GSpecies, HSpecies, ISpecies, JSpecies, KSpecies, LSpecies, MSpecies, NSpecies, OSpecies, PSpecies, QSpecies, RSpecies, SSpecies, TSpecies, USpecies, VSpecies, XSpecies, YSpecies, ZSpecies, AASpecies, ABSpecies, ACSpecies, ADSpecies, AESpecies)

speciesByYear <- data.frame(YEAR, speciesChrono)

# look at it graphically

par(mfrow=c(1,1))
plot(speciesByYear$YEAR, speciesByYear$speciesChrono, xlab = "Year", ylab = "Species Number")

```

# Discussion

# Reproducibility

The unmanipulated dataset and analysis are hosted on a Github repository (linked in the Methods section), along with the manipulated dataset and corresponding code, figures, license information and a Docker file. The dataset was not called for using an API, making it necessary to download externally and import into R. The original directions are documented in the analysis (on Github) and the unmanipulated dataset can also be downloaded from the Github repository. As mentioned earlier, the original dataset had some issues that had to be resolved prior to analysis. Abundances of “zero” were not entered into the dataset, leaving that year out of the dataset entirely. Several species were also left out, but using the orignal study were returned to the data set. An image from the original study is in the analysis folder on the Github repository for easy reference (Morris, 1991). Once these were returned the dataset, further manipulation and analysis was possible. The analysis, an R markdown file, details how to download the dataset and shows the code for manipulation in order to minimize any reproducibility issues and show transparency.

# Conclusions

This project was inherently limited by the structure and lack of completeness of the data set obtained. Perhaps the most important conclusion of this project is the forethought required in how data is collected and databased for future use. Morris, the original researcher, collected this information in a time when these considerations did not need to be taken, as open source data sets were certainly not on the average researcher's mind. While this study is limited in its analytical ability, the author does not intend to masquerade the results as such. Much of the ecological information needed in order to determine the most important factors in *Odonata* diversity (vegetation, pond area, scrub management) is not available. Therefore, the findings of the original study cannot be expanded upon in an ecological sense. It also showcases the difficulty in finding representative and complete data sets, with the admission that this data set might not have been the best choice, at least not on its own.

# References

Dornela M, Antao LH, Moyes F, Bates A. E., *et al.* BioTIME: A database of biodiversity time series for the Anthropocene. *Global Ecol Biogeogr.* 2018; 27:760-786. https://doi.org/10.1111/geb.12729

Kalkman, V. J., Clausnitzer V., Dijkstra K. B., Orr A. G., Paulson D. R., van Tol J. (2008) Global diversity of dragonflies (Odonata) in freshwater. *Hydrobiologia*, **595**, 351-363.

Moore, N. W. (1991) The development of dragonfly communities and the consequences of territorial behavior: a 27 year study on small ponds at Woodwalton Fen, Cambridgeshire, United Kingdom. *Odonatologica*, **20**(2), 203-231.

Waage, J. K. (1979) Dual function of the damselfly penis: Sperm removal and transfer. *Science*, **203**:4383, 916-918.

The Wildlife Trusts (n.d.) About the Great Fen. https://www.greatfen.org.uk/about-great-fen